---
title: "Alewife Telemetry - Lake Ontario - Nowcast Temperatures"
date: "Updated: `r format(Sys.time(), '%Y-%m-%d %H:%M:%S')`"
---
# Purpose
This site is designed to provide quick updates on projected water temperatures given current conditions
from [NOAA's GLERL website](https://www.glerl.noaa.gov/res/glcfs/ncast.php?lake=ont) for specific areas of Lake Ontario. 

This site will automatically rebuild every 12 hr and will pull the 
latest projection. 

# Area of interest 
Below is a interactive map with the area and corresponding 'nodes' or locations 
in relation to the projected data. 

```{r, message=FALSE, echo = FALSE}
#| label: load pkgs
# ---- load pkgs ----
{
  library(bslib)
  library(curl)
  library(dplyr)
  library(ggplot2)
  library(here)
  library(lubridate)
  library(mapview)
  library(plotly)
  library(qs)
  library(sf)
  library(stringr)
  library(tidync)
}
```


```{r, echo=FALSE}
#| label: load bnd 

array_bnd <- qread(here('data',
                        'array_bnd.qs'))
```


```{r, echo=FALSE, message = FALSE}
#| label: get url to file 
# the base url 

base_url <- 'https://www.glerl.noaa.gov/emf/ont/ncast/'
# first we grab the current time to download the latest push 
current_time <- paste0(format(Sys.time(), "%m%d"),
                       if_else(as.integer(format(Sys.time(), "%H")) < 12,
                               true = "00", 
                               false = "12"))

# we need to add the last strip to the end of the current time 
file_name <- sprintf('%s_0001.nc', current_time)

# paste teh base url and file_name together to get correct file 
url <- paste(base_url, file_name, sep = '')
```

```{r, echo=FALSE, message=FALSE}
#| label: downloaded lastest netcdf 
# use curl to donwlaod 
curl_download(
  url = url,
  destfile = file_name,
  handle = curl::new_handle(followlocation = TRUE),
  mode = "wb", 
)
# ---- load nc file 
nc <- tidync(file_name)
```

```{r, echo=FALSE, message=FALSE}
#| label: grab locations 
locs <- nc %>%
  activate("D1") %>%
  hyper_tibble() %>% 
  mutate(
    lon = lon - 360
  ) %>%
  select(
    node, lon, lat, h
  )
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#| label: convert to sf 
locs_sf <- locs %>%
  st_as_sf(coords = c("lon", "lat"),
           crs = 4326)

locs_sf_select <- st_intersection(locs_sf, array_bnd)
```

```{r, echo=FALSE, warning=FALSE}
#| label: plot map  
mapview(array_bnd, legend = FALSE) + 
  mapview(locs_sf_select, zcol = "glatos_array", 
          layer.name = "Glatos Array") 
```

```{r, echo=FALSE}
#| label: drop sf 
locs_select <- locs_sf_select %>% 
  mutate(lon = st_coordinates(.)[,"X"],
         lat = st_coordinates(.)[,"Y"]
  )  %>%
  st_drop_geometry() %>%
  as_tibble()
```


```{r, echo=FALSE}
#| label: grab temps
temps <- nc %>%
  activate("D1,D2,D5") %>%
  hyper_tibble() %>% 
  filter(node %in% unique(locs_select$node)) %>% 
  mutate(
     time = str_replace(time, "T", " ") %>%
      ymd_hms()
  )
```


```{r, echo=FALSE, message = FALSE}
#| label: join temps with locations 
temps_joined <- temps %>%
  left_join(
    locs_select
  )
```

```{r, echo = FALSE}
#| label: add sig and depths 

time_n <- length(unique(temps_joined$time))

temps_joined <- temps_joined %>% 
  arrange(node, time) %>% 
  mutate(
    sig = rep(rep(seq(0.025, 0.975, 0.05), time_n), nrow(locs_select)), 
    depth = h * sig
  )
```

# Predicted thermal profiles

Below are predicted thermal profiles for nodes `3383`, `4512`, and `5088` which are the center of the array.

```{r, echo = FALSE, warning=FALSE}
#| label: create thermal profiles in ggplot 
p <- ggplot(data = temps_joined %>% 
         filter(node %in% c("3383", "4512", "5088")),
            aes(x = time, y = depth,
                fill = temp)) +
  geom_raster() +
  scale_fill_viridis_c(name = "Temperature (°C)",
                       option = "H", 
                       breaks = seq(5, 20, 2.5)) +
  scale_y_reverse(breaks = seq(0, 70, 5)) +
  facet_wrap(~ glatos_array, scale = "free_y") +
  theme_bw(base_size = 15) +
  theme(
    strip.background = element_blank(), 
    plot.title = element_text(hjust = 0.5)
  ) +
  labs(
    x = "Date",
    y = "Depth (m)", 
    title = paste("Nowcast end date:", current_time, sep = " ")
  )
```

```{r, echo=FALSE, warning = FALSE}
#| label: nmake them a widget with plotly
#| fig-width: 10
#| fig-height: 7

ggplotly(p) %>% 
  card(full_screen = TRUE)
```

